% *** Théorie
% *** Exemples d'applications

\section{Methods}
\label{sec:definitions}

The HLDA is an extension of the multiclass LDA.
It's goal is to allow heteroscediscity: The classes respective Covariance matrix can be different.
First, we will explain the bases of the LDA.
Then we will analyze how the HLDA is an extension of these bases.

\subsection{Notations}

The following notations are used throughout the rest of the document:

\begin{itemize*}
  \item Let $X_i$, $i \in \llbracket 1 \ldots N \rrbracket$ be a $n$-dimensional vector corresponding to a sample.
  \item We consider $J$ classes and refer to a class by $C_j$, $j \in \llbracket 1 \ldots J \rrbracket$
  \item The number of samples in a class is given by $N_j$, $j \in \llbracket 1 \ldots J \rrbracket$
  \item The notation $X_i^{(j)}$, $j \in \llbracket 1 \ldots J \rrbracket$, $i \in \llbracket 1 \ldots N_j \rrbracket$
    represents the ith sample of the jth class
  \item For each classes, the mean $\mu_j$ and the covariance $\Sigma_j$ are defined as:
    $$\mu_j = \frac{1}{N_j}\sum\limits_{i = 1}^{N_j} X_i^{(j)}$$
    $$\Sigma_j = \frac{1}{N_j}\sum\limits_{i = 1}^{N_j} (X_i^{(j)} - \mu_j)(X_i^{(j)} - \mu_j)^t$$
  \item Finally the mean and covariance of the full dataset are defined as:
    $$\mu = \frac{1}{N}\sum\limits_{i = 1}^{N} X_i$$
    $$\Sigma = \frac{1}{N}\sum\limits_{i = 1}^{N_j} (X_i - \mu)(X_i - \mu)^t$$
\end{itemize*}

\subsection{Multiclass LDA}

\subsubsection{Purpose}

The main goal of the multiclass LDA is to find a predictor that will allow to find the class that best correspond to a sample X.
In order to do that, we need to find a separator and each class will be projected on it.
The main goal is to find the best separator, so for a sample X, it will be easy to find which class
best correspond to X.
In a second time, it is possible to perform a dimension reduction on the separator.

To summarize, the LDA has two objectives:
- Find a separator that will allow a sample to be classified.
- Perform a dimension reduction to classify faster.

Now we will explain how the multiclass LDA works. 

\subsubsection{Methodology}

We want to find the class that best correspond to a vector x.
In order to easily differentiate classes, we need:

\begin{enumerate}
\item Each projected sample of the same class should be very close: we will call this the within-class ratio $S_W$.
\item The distance between each class data should be very large: we will call this the between-class ratio $S_B$.
\end{enumerate}

We will try to find the smallest within-class ratio and the biggest between-class ratio in order to best isolate each class.

So the function that we want to maximize is:
$$S = \frac{w^t S_B w}{w^t S_W w}$$
with:
\begin{itemize*}
  \item $S$ the ratio to maximize
  \item $S_B$ the between-class ratio
  \item $S_W$ the within-class ratio
  \item $w$ the projection matrix
\end{itemize*}

The expressions of $S_B$ and $S_W$ for the multiclass LDA are:
$$S_B = \frac{1}{C} * \sum\limits_{i = 1}^C (\mu_i - \mu)(\mu_i - \mu)^t$$
$$S_W = \frac{1}{C} * \sum\limits_{i = 1}^C \sum\limits_{\forall x \in C_i} (x - \mu_i)(x - \mu_i)^t$$
With:
\begin{itemize*}
  \item $\mu$ the mean of the class means
  \item $\mu_i$ the mean of the ith class
\end{itemize*}

For a multiclass LDA, the best projection matrix for the ith dimension satisfy the equation:
$$(S_B - \lambda_i S_W)w_{i}^* = 0$$

This means that the projections with the maximum class separability information are the eigenvectors corresponding to the largest eigenvalues of $S_{W}^{-1} S_B$.

\subsection{HLDA}

\subsubsection{Differences between LDA and HLDA}

HLDA (Heteroscedastic Linear Discriminant Analysis) is another method to find a linear transformation mapping the problem space into a smaller one.

The HLDA finds the projection matrix A such that:
$$y = Ax$$
with y a $p$ dimensions variable and $p < n$.\\

The purpose of the HLDA method is to release the homoscedasticiy constraint of the LDA, allowing different variances between classes. The two assumptions made by the method are:

\begin{enumerate}
  \item The class distributions are $n$-dimensional Gaussian distributions.
  \item This space of size $n$ can be split into two subspaces of size $p$
    and $n - p$ (with $p < n$). The first one is different for all classes and contains meaningful information while the second one is the same for the whole dataset and contains only noise.
\end{enumerate}

The projection matrix A can then be rewritten following those assumptions:
$$A = \left [
  \begin{array}{c}
    A_p\\
    A_{n-p}
  \end{array}
\right ]
$$

it is also possible to find an expression for the mean and variance of each classes:
$$\mu_j =
\left [
  \begin{array}{c}
    \mu_{j,p}\\
    \mu_{0,(n-p)}
  \end{array}
\right ]
$$

$$\Sigma_j =
\left [
  \begin{array}{cc}
    \Sigma_j^p & 0\\
    0 & \Sigma_0^{(n-p)}\\
  \end{array}
\right ]
$$
Where $\mu_0$ and $\Sigma_0$ are supposed to be common to all classes.

Based on those definition, it is possible to define a model representing the new dataset.

\subsubsection{HLDA gaussian Model}

The Gaussian probability density function for this model is:
$$P(y_i) = \frac{\mid \theta \mid}{\sqrt{(2\pi)^n \mid \Sigma_{g(i)} \mid}}
    e^{-\frac{1}{2} (y_i - \mu_{g(i)})^t \Sigma_{g(i)}^{-1} (y_i - \mu_{g(i)})}$$
With $g(i)$ the function associating each sample with its label.

The variables of the model are:
\begin{itemize}
  \item $A$: the projection matrix
  \item $\mu_j$ the mean of class $j$
  \item $\Sigma_j$ the covariance of class $j$
\end{itemize}

Note that $\mu_j$ and $\Sigma_j$ can depend on $A$ thus no analytical solution can be found (in contrast to the LDA method). In order to find estimators for those parameters Kumar\cite{kumar.1997} uses the maximum log-likelihood method and find two different solutions depending on the type of $\Sigma_j^p$.

\subsubsection{Full-rank $\Sigma_j^p$}

%gary

\subsubsection{Diagonal $\Sigma_j^p$}

%gary

Note that when considering equals $\Sigma_j$ are the same as the LDA.

\subsubsection{Computing $\hat \theta$}

%gary

%%   L(\{x_i\}; A; \{\mu_j, \Sigma_j\}) & = & \sum\limits_{j=1}^{C}\sum_{i=1}{N_j} log p_j x_i^{(j)}\\
%%                                      & = & \sum\limits_{j=1}^{C}\sum_{i=1}{N_j} [\frac{1}{2} log (\frac{(\mid A \mid)^2}{(2\pi)^n \product_{k=1}^n
%% \end{array}$$
